---
output: rmarkdown::github_document
---

### #TidyTuesday 10-09-18 Week 28

This week's data explores voter turnout across the US from 1980 through 2012. The data can be found [here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2018-10-09).

Today we're going to do a boxplot visualizing percent of voter turnout by region for presidential election years.

First, let's read in our data using `read_csv()`. We'll use `head()` to get a look at what we're working with.

```{r}
library(tidyverse)

data <- read_csv("voter_turnout.csv")

head(data)
```

I found a list [here](https://github.com/cphalpert/census-regions/blob/master/us%20census%20bureau%20regions%20and%20divisions.csv) that has each of the states & it's corresponding region & division. For clarity of the plot later on, I chose to use region instead of division. There are 4 regions & 9 divisions. 4 regions for each year will be easier to plot than 9 divisions for each year. Let's read in the region/state file & see what we're working with.

```{r}

region_list <- read_csv("state.region.list.csv")

head(region_list)
```

Before we start manipulating our data frame, we need to establish which years a presidential election was held. We can use the command below to create a list of presidential election years.

```{r}

president_years <- seq(1980, 2012, 4)
```

Now let's work on cleaning up our data frame. First, we're going to use `filter()` to keep only the years when a presidential election was held. We can use `%in%` to filter the data frame using the list we created in the last step. Next, we use `select()` to get the columns we need. `mutate()` is used to find the percent of eligible voters that actually voted. We can use the same `mutate()` call to rename our "state" column. This way it matches the name of the column in the region/state data frame.

```{r}
voter <- data %>%
  filter(year %in% president_years) %>%
  select(year, state, votes, eligible_voters) %>%
  mutate(pct = votes/eligible_voters, State = state)

head(voter)
```

Ugh! Look at those NAs. Don't worry! We'll get rid of them in just a moment.

 First, let's join our two data frames. Here, we're going to use `left_join` to combine the voter data frame & the region/state data frame. We specify the voter data frame first because we want the region/state to fill in on each appropriate line of the voter data frame. Then we use `na.omit()` to remove all lines with an NA from the data frame. Finally, we use `select()` to keep only the columns we need for our plot.

```{r}
final_voter <- left_join(voter, region_list, by = "State") %>%
 na.omit() %>%
 select(year, Region, pct) %>%
 group_by(year, Region)

head(final_voter)
```

Let's get to the plot! We're going to start with a basic box plot. To do this we use `geom_boxplot()`. We specify the year on the x-axis, the percent on the y-axis, & the fill as Region. All three things are specified within the `aes()` call in `geom_boxplot()`.

Interesting side note - For the box plot to work, one of the axes needs to be a categorical variable. We would think of year as a categorical variable. R, however, does not think of year the same way we do because it's a numeric value. By adding `factor()` into the `aes()` call, we can convince R that year is indeed a categorical variable.

```{r, fig.height = 7.5}
ggplot(final_voter) +
  geom_boxplot(aes(x = factor(year), y = pct, fill = Region))
```

Looks good. We have what we would expect. The first thing I would like to do is use `coord_flip()` to flip the graphic on it's side. I think it looks better that way. I also want to move the legend to the bottom to give the plot a bit more room. This can be done in the `theme()` call.

```{r, fig.height = 7.5}
ggplot(final_voter) +
  geom_boxplot(aes(x = factor(year), y = pct, fill = Region)) +
  coord_flip() +
  theme(legend.position = "bottom")
```

Next, let's change up the color scheme a bit. We can do this using `scale_fill_manual()` and creating a vector with our 4 hex codes in it.

Let's also remove the back panel & add in some dashed grid lines. Both of these can be done from the `theme()` call. To remove the background panel, we use `panel.background = element_blank()`. Changing the grid lines is a bit more complicated. We use `panel.grid.major.x = element_line(colour = "black", linetype = "dashed", size = 0.1)`. You can play with the line size & type until it looks how you would like.

```{r, fig.height = 7.5}
ggplot(final_voter) +
  geom_boxplot(aes(x = factor(year), y = pct, fill = Region)) +
  coord_flip() +
  theme(legend.position = "bottom",
    panel.background = element_blank(),
    panel.grid.major.x = element_line(colour = "black", linetype = "dashed", size = 0.1)) +
  scale_fill_manual(values = c("#2a9d8f", "#e9c46a", "#f4a261", "#e76f51"))
```

I would like to remove the titles from each axis & change the font size on the labels. As before, both of these things can be done from the `theme()` call. `axis.title = element_blank()` removes the title from both axes. `axis.text = element_text(size = 12)` changes the font size of the labels of both axes.

I would also like to change the scale on the x-axis to percent. I used `scale_y_continuous(labels = scales::percent)` to make that happen. I used `scale_y_continuous()` because the x-axis in our plot was the original y-axis before we used `coord_flip()`.

Last of all, I want to remove all the tick marks from both axes. I do this by adding `axis.ticks = element_blank()` to the `theme()` call.

```{r, fig.height = 7.5}
ggplot(final_voter) +
  geom_boxplot(aes(x = factor(year), y = pct, fill = Region)) +
  coord_flip() +
  theme(legend.position = "bottom",
    panel.background = element_blank(),
    panel.grid.major.x = element_line(colour = "black", linetype = "dashed", size = 0.1),
    axis.title = element_blank(),
    axis.text = element_text(size = 12),
    axis.ticks = element_blank()) +
  scale_fill_manual(values = c("#2a9d8f", "#e9c46a", "#f4a261", "#e76f51")) +
  scale_y_continuous(labels = scales::percent)
```

I would call this a finished plot!

There are a few other things I wanted to add about boxplots that don't really work with the data set we have here.

If you want to add the actual points onto the boxplot, you could add `geom_jitter()` to your plot. You could add `geom_jitter(aes(x = factor(year), y = pct))` & this would plot all the data points for you.

Also, you can add a notch into the boxplot at the median line. To do this would add `notch = TRUE` to the `geom_boxplot()` call. Example - `geom_boxplot(aes(x = factor(year), y = pct, fill = Region), notch = TRUE)`.

Also, you can experiment with varidable width boxplots. This is also done from the `geom_boxplot()` call. Example - `geom_boxplot(aes(x = factor(year), y = pct, fill = Region), varwidth = TRUE)`.

If you have any questions or comments, please feel free to reach out to me via [Twitter](https://twitter.com/sapo83).
