---
output: rmarkdown::github_document
---

### #TidyTuesday 1-1-19 Week 1

It's the beginning of a new year & I'm already behind! What's new?

This week's data is a peek into the statistics of last year's #TidyTuesday & #Rstats tweets.
I'm going to start off with an analysis of the number of Twitter users who tweeted aboout #Rstats broken down by country. All the data used in this analysis can be found [here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-01-01).

Let's start by loading the tidyverse library & our data. This data is provided in .rds format. A quick google search shows that `readRDS()` from base R can be used to read in this type of data. Then I'll use `head()` to take a quick look at the data.

```{r setup}
library(tidyverse)

data <- readRDS("rstats_tweets.rds")

head(data)
```

That's a lot of columns!!! I'm going to start by selecting the columns I need using `select()`. Since the same person may have tweeted multiple times, I need to save only one line where their user name & country is recorded. I can use `distinct()` to keep only the unique lines. I'm going to remove all users with an `NA` for the country.

```{r userDF}
user_df <- data %>%
  select(screen_name, country) %>%
  distinct() %>%
  drop_na()

user_df
```

A quick look at this data frame shows that some users have multiple countries reported.
To take this into account, we're going to use the number of rows in this data frame to calculate our percentages. To find the number of rows in the data frame, I can use `nrow()`. I'm going to save the number of rows to a variable (rowNum).

```{r numRows}
rowNum <- nrow(user_df)

rowNum
```

Now let's get the counts for each country. First, we group the data frame by country using `group_by()`. Then we use `summarise()` to get the counts. Then, we'll order the data frame by descending count.

```{r count}
count_df <- user_df %>%
  group_by(country) %>%
  summarise(count = n()) %>%
  arrange(-count)

count_df
```

Next I'll create the percent data frame. First, we'll keep the top 10 counts. For this, I'll use `top_n()`. If you wanted to get the bottom 10 values, you could use `top_n(-10)`. I'll use mutate to calculate the percentage using the row number variable we created earlier.

```{r percent}
pct_df <- count_df %>%
  top_n(10) %>%
  mutate(pct = (count/rowNum) * 100)

pct_df
```

Next, I'm going to plot the data. I'm going to use a bar chart.

```{r plot1}
ggplot(pct_df) +
  geom_col(aes(x = country, y = pct))
```

Not bad! Now let's make it pretty. I'm going to add some color to the bars by adding `fill = ` to the `aes()` call. I'm going to use `coord_flip()` to flip the plot on it's side. This will make the countries easier to read.

```{r plot2}
ggplot(pct_df) +
  geom_col(aes(x = country, y = pct, fill = country)) +
  coord_flip()
```

I want to reorder the (old) x-axis by percent, highest to lowest. I can do this by using `reorder()` in the `aes()` call. I'm also going to remove the legend using the `theme()` call.

```{r plot3}
ggplot(pct_df) +
  geom_col(aes(x = reorder(country, pct), y = pct, fill = country)) +
  coord_flip() +
  theme(legend.position = "none")
```

Then I want to add precent labels to the bars so the reader can identify the acutal values. I can do this by adding a `geom_text()` call. I'm going to use `round()` to format them to two decimal places. I'll use paste to add a "%" symbol as well. Lastly, I'll use `hjust` to change the justification of the labels. I would like them at the end of the bar.

```{r plot4}
ggplot(pct_df) +
  geom_col(aes(x = reorder(country, pct), y = pct, fill = country)) +
  geom_text(aes(x = reorder(country, pct), y = pct, label = paste(round(pct, 2), "%")), hjust = -0.25) +
  coord_flip() +
  theme(legend.position = "none")
```

Oops! Now the label can't be fully shown in the plot area. This can be fixed by changing the limits on the plot area. This can be done by simple trial & error. As I make more & more plots, I've tried to find a more progrommatic way of doing this. We're going to use a variable to determine what the plot area should be.

I'm going to create a variable named "max" that holds the maximum precentage in the data frame.
Then `ylim()` can be used to change the y-axis limits. I like this way because if the data in the data frame changes then the y-axis limit changes along with it.

```{r plot5}
max <- max(pct_df$pct)

ggplot(pct_df) +
  geom_col(aes(x = reorder(country, pct), y = pct, fill = country)) +
  geom_text(aes(x = reorder(country, pct), y = pct, label = paste(round(pct, 2), "%")), hjust = -0.25) +
  coord_flip() +
  ylim(0, max * 1.10) +
  theme(legend.position = "none")
```

Let's do a final clean up! I'm going to add a title using `ggtitle()`. Using the `theme()` call, I'm going to remove the background panel, remove the title & tick marks from both axes, change the font size on the y-axis, remove the text from the x-axis, & center the title on the plot.

```{r plotFin}
max <- max(pct_df$pct)

ggplot(pct_df) +
  geom_col(aes(x = reorder(country, pct), y = pct, fill = country)) +
  geom_text(aes(x = reorder(country, pct), y = pct, label = paste(round(pct, 2), "%")), hjust = -0.25) +
  coord_flip() +
  ylim(0, max * 1.10) +
  ggtitle("#Rstats Tweets By Country") +
  theme(legend.position = "none",
    panel.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_blank(),
    plot.title = element_text(hjust = 0.5))
```

That looks great! If you have any feedback/questions, please feel free to create an issue or pull requet. You can always connect with me on [Twitter](https://twitter.com/sapo83).
