---
output: rmarkdown::github_document
---

### #TidyTuesday 8-28-18

This week's data is NFL statistics. I only know a little bit about football so the visualization I did may not be super informative, but half the learning is in the process. I got the general idea from the [article](https://www.eldo.co/nfl-rushing-and-passing-in-four-charts.html) included with the data.  Let's get started!

First, we want to load the R package we are going to use. Since it's #TidyTuesday, we'll be using the tidyverse. Next step is to read in our data. The .csv file can be downloaded [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2018-08-28/nfl_2010-2017.csv).

```{r}
library(tidyverse)

data <- read_csv("nfl_2000-2017.csv")

data
```

This gives us a general idea of what is included in this data set. I decided to focus on fumbles. There are three types of fumbles: rush fumbles, receiving fumbles, & pass fumbles. There are also multiple players listed per team. We need to create a data frame that totals each type of fumble for all the players on that team for each year.

First, we use select() to select our columns of interest & remove all the other columns. There are quite a few NAs in this data. For this analysis, I set all NAs to 0 using mutate_if(). Next, we group the rows by team & year using group_by(). Then the summarise() function allows us to total each type of fumble for all the players listed on that team in that year.

```{r}
sum_team_year <- data %>%
  select(name, team, game_year, game_week, rush_fumbles, rec_fumbles, pass_fumbles) %>%
  mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .))) %>%
  group_by(team, game_year) %>%
  summarise(rush_fum = sum(rush_fumbles), rec_fum = sum(rec_fumbles), pass_fum = sum(pass_fumbles))

sum_team_year
```

In the resulting data frame, you can see each team has one row for each year with the appropriate number of fumbles.

Next, we want to find the number of fumbles for all the teams each year. We start with the previous data frame (sum_team_year). First, we use the ungroup() to remove our previous grouping. Then we can drop the team column because it is no longer needed. We regroup the data frame by year using group_by(). Then we use the summarise function to get the total number of each type of fumble for all the teams each year.

```{r}
sum_year <- sum_team_year %>%
  ungroup() %>%
  select(-team) %>%
  group_by(game_year) %>%
  summarise(rush_fum_total = sum(rush_fum), rec_fum_total = sum(rec_fum), pass_fum_total = sum(pass_fum))

sum_year
```

This leaves us with a data frame with 18 years. This makes sense becuase we have 18 years in our dataset. Little checks like this along the way can help us make sure we're on the right track!

Next, I want to find how many fumbles there were per team for each year. To do this, I first need to determine the number of teams per year. A quick Google search tells me that there are not necessarily the same number of teams every year. Instead, I'll use our data to make a new data frame with the number of teams per year.

We can use the original data() object to make this data frame. First, we select the columns we are interested in. Since there are multiple players per team & I only wanted each team counted once, I used distinct() to remove all the duplicate rows. Then I used select() to remove the team column. I grouped the remaining data by year using group_by(). I used count() to count the number of times each year appeared in the data frame. Each year will appear once for each team that existed that year.

I used rename() to change the name of the "number of teams" column to make things easier down the road. When you use rename(), the correct usage is (new_col_name = old_col_name).


```{r}
num_teams <- data %>%
  select(team, game_year) %>%
  distinct() %>%
  select(-team) %>%
  group_by(game_year) %>%
  count(game_year) %>%
  rename(num_team = n)

num_teams
```

Then we need to join this data frame with the number of teams per year with the yearly fumble data. There are a lot of different types of joins. I'm not going to go too deep into that today. For this particular example, I used a full join. A full join uses all rows in both data frames. I think this is a good way to catch random data you may not realize is in your data frame.

The full_join() function takes three arguments: the first data frame, the second data frame, & the column to join them by. Then I used mutate() to determine how many fumbles per team there were for each year. I used select() to retain only the columns I'm interested it.

If you have lots of columns with the same naming pattern or word, there are a couple of helper functions you can use in your select() statement.

  * starts_with()
    * finds column names that start with a certain string
  * ends_with()
    * finds column names that end with a certain string
  * contains()
    * finds column names that contain that string.

For this example, I used contains() to keep all the columns that have "per_team" in their name.

After using select to get the columns I need, I used melt from the reshape2 package to convert the data frame from wide to long.

```{r}
total_df <- full_join(sum_year, num_teams, by="game_year") %>%
  mutate(rush_fum_per_team = (rush_fum_total/num_team), rec_fum_per_team = (rec_fum_total/num_team), pass_fum_per_team = (pass_fum_total/num_team)) %>%
  select(game_year, contains("per_team")) %>%
  reshape2::melt(id.vars="game_year")

total_df
```

Now we're ready to plot our data! For this, I decided to use a line plot to vizualize the three different groups.

```{r}
ggplot(total_df) +
  geom_line(aes(game_year, value, group=variable, color=variable))
```

Let's add some points to make the years stand out a bit more!

```{r}
ggplot(total_df) +
  geom_line(aes(game_year, value, group = variable, color = variable)) +
  geom_point(aes(game_year, value, group = variable, color = variable))

```

Now let's clean it up a bit! First, I would like to change the labels in the legend. To do this, I need to rename the factors. This is the first line added into the code below. Next, I want to relabel the axes so they make more sense. I also used theme() to move the legend to the bottom of the plot & remove the title.

```{r}
levels(total_df$variable) <- c('Rush Fumbles', 'Receving Fumbles', 'Pass Fumbles')

ggplot(total_df) +
  geom_line(aes(game_year, value, group = variable, color = variable)) +
  geom_point(aes(game_year, value, group = variable, color = variable)) +
  xlab("Year") +
  ylab("Average Per Team") +
  theme(legend.position = "bottom", legend.title=element_blank())

```

Next I want to change the color for each group. I can do this using scale_color_manual(). Note: If you used fill to color whatever type of graph you have, changing the color would use scale_fill_manual().
I added to the theme() function to change the major grid lines to dotted gray lines. Then I removed the minor grid, the panel border & the panel background. I also bumped up the font size for the axis title, tick mark labels, & the soon to be added title. I used scale_x_continuous() & the scales package to add more tick marks. Last of all, I added a title to the plot.

```{r}
levels(total_df$variable) <- c('Rush Fumbles', 'Receving Fumbles', 'Pass Fumbles')

ggplot(total_df) +
  geom_line(aes(game_year, value, group = variable, color = variable)) +
  geom_point(aes(game_year, value, group = variable, color = variable)) +
  xlab("Year") +
  ylab("Average Per Team") +
  theme(legend.position = "bottom",
    legend.title=element_blank(),
    panel.grid.major = element_line(colour = 'azure3', linetype = 'dashed'),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 16) ,
    plot.title = element_text(size = 20)) +
  scale_x_continuous(breaks=scales::pretty_breaks(10)) +
  scale_color_manual(values=c("orange", "blue", "red")) +
  ggtitle("Average Number of Fumbles Per Team")

```

Something is going on with the data for 2005 through 2008. There are a lot of NAs in the fumbles data for these four years. I'm not sure why this is. All the NAs were set to 0 so the averages end up being veyr low. Any insight would be appreciated. You can reach out to me via Twitter at @sapo83 if you have any ideas! Also if you have any other thoughts or insights about this tutorial. 
