---
output: rmarkdown::github_document
---

### #TidyTuesday 05-14-2019 Week 20

This week's data looks at Nobel Prize winners & the associated metadata. I decided to focus on the birth country of each winner. I thought this was a good chance to calculate some percents & used a stacked bar chart.

Let's start by loading the tidyverse library & reading in our data.

```{r setup}
library(tidyverse)

nobel_winners <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-14/nobel_winners.csv")

head(nobel_winners)
```

Data looks ok. Let's start by finding the total number of Nobel Prizes awarded by birth country. First, I used `select()` to keep only the birth country. I dropped all NAs to make life easier! Then I grouped by birth country & used `summarise()` to get the total for each country. I used `top_n()` to keep the 12 countries with the highest totals. Last of all, I saved these 12 countries into a list for later use.

```{r cleanup}
count_list <- nobel_winners %>%
  select(birth_country) %>%
  drop_na() %>%
  group_by(birth_country) %>%
  summarise(total = n()) %>%
  top_n(12, total)

country_list <- count_list$birth_country

head(country_list)
```

Next, I'm going to filter the data & calculate our percents. I used `select()` to get the columnds I needed. I used `filter()` to keep only the countries in our top 12 list. I grouped by birth country & category & used `summarise()` to get the count of each type of Nobel Prize for each country. I used right join to add the column with the total number of Prizes for each country. Finally, I used `mutate()` to calculate the percent for each type of Nobel Prize.

```{r filter}
filter_df <- nobel_winners %>%
  select(birth_country, category) %>%
  filter(birth_country %in% country_list) %>%
  group_by(birth_country, category) %>%
  summarise(count = n()) %>%
  right_join(count_list) %>%
  mutate(pct = (count/total) * 100)

head(filter_df)
```

For the plot, I want to change the order of countries to reflect the total numebr of Nobel Prizes earned. I used `reorder()` to set this order.

```{r plot}
filter_df$birth_country <- reorder(filter_df$birth_country, filter_df$total)
```

I'm going to work my way through this line by line.

`geom_bar()` sets the fill by type of Nobel Prize. I used `fct_rev()` to get alphabetical order from left to right.

`geom_text()` adds the total number of Prizes on the right hand side of the bar.

`coord_flip()` changes the plot from vertical bars to horizontal bars.

`scale_x_discrete()` sets the length of the word wrap on the y-axis of the final plot.

`scale_fill_manual()` sets the colors for each category.

`guides()` reverses the order of the legend so it matches the order of the plot.

`ggtitle()` sets the title of the plot.

`theme()` does the following:

  - move legend to bottom of plot
  - remove legend title
  - remove axes title
  - make axis text font larger
  - remove panel background
  - remove axis ticks
  - format plot title

```{r fin_plot}
ggplot(filter_df) +
  geom_bar(aes(x = birth_country, y = pct, fill = fct_rev(category)), stat = "identity") +
  geom_text(aes(x = birth_country, y = 105, label = total)) +
  coord_flip() +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 15)) +
  scale_fill_manual(values = c("#5EBD3E", "#FFB900", "#F78200", "#E23838", "#973999", "#009CDF")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 6)) +
  ggtitle("Nobel Prizes Earned by Type (%) & Total \nBased on Birth Country") +
  theme(legend.position = "bottom",
    legend.title = element_blank(),
    axis.title = element_blank(),
    axis.text = element_text(size = 14),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 20, hjust = 0.5))
```

Looks good! If you have any feedback/questions, please feel free to create an issue or pull request. You can always connect with me on [Twitter](https://twitter.com/sapo83).
