---
output: rmarkdown::github_document
---

### #TidyTuesday 9-18-18 Week 25

This week's data is US airport [data](https://github.com/rfordatascience/tidytuesday/blob/master/data/2018-09-18/us-airports.csv).

Since I live in Illinois, I've decided to make a state map showing all the airports in Illinois & their yearly rank.

First, let's start by loading the libraries we need & the data file. Then, let's use head to take a quick peek at our data.

```{r  setup}
library(tidyverse)
library(ggmap)
library(maps)

data <- read_csv("us-airports.csv", col_names = TRUE)

head(data)
```

To complete this map, we're going to need the latitude & longitude for each city in our data set. I found a csv file with the longitude & latitude for all the cities [here](https://simplemaps.com/data/us-cities).

First, we read in the data. Then we use select() to pick the columns we need. Next, we filter the data set by the state of interest (IL). To make joining the two data frames easier later, I used mutate to add a column with the state. Then I used select to drop the state_id column. Once again, we use head to look at our data & make sure everything still looks good.

```{r lat/long df}
long_lat_df <- read_csv("us.cities.csv")

long_lat_il_df <- long_lat_df %>%
  select(city, state_id, lat, lng) %>%
  filter(state_id == "IL") %>%
  mutate(state = state_id) %>%
  select(-state_id)

head(long_lat_il_df)
```

Now that our longitude & latitude data frame is set, let's go back to our original data set.

First, we use select() to get the columns we need for this analysis. Then we filter by state. I used to distinct() to make sure that we only have one record for each airport.

Next, I used mutate_if to clean up some of the city names. In the airport data, some of the cities are a bit strange. Using str_replace_all, I can search for a pattern & replace it with the pattern of my choice.

Last of all, I used mutate to remove everything from the city names in parentheses.

```{r IL df}
IL_airport_df <- data %>%
  select(yearly_rank, state, city, airport_name) %>%
  filter(state == "IL") %>%
  distinct() %>%
  mutate_if(is.character, str_replace_all, pattern = "Bloomington-Normal Airport", replacement = "Bloomington") %>%
  mutate_if(is.character, str_replace_all, pattern = "De Kalb", replacement = "DeKalb") %>%
  mutate_if(is.character, str_replace_all, pattern = "DuPage", replacement = "West Chicago") %>%
  mutate_if(is.character, str_replace_all, pattern = "Romeo", replacement = "Romeoville") %>%
  mutate(city = trimws(str_replace(city, "\\(.*?\\)", "")))

head(IL_airport_df)
```

Now that our airport data is cleaned up, we can merge it with the longitude/latitude data from before. Here I used left_join(). The first data frame is the Illinois airport data frame. Order matters here. When using left join, the join function will keep all the rows in the first data frame listed in the command.

There is also a "group" value we need for plotting the map. A quick look at the data frame to plot the map shows the group for Illinois is "12". I used mutate to add a group column with a value of "12".

```{r merge two df}
IL_df <- left_join(IL_airport_df, long_lat_il_df, by = c("state", "city")) %>%
  mutate(group = "12")

head(IL_df)
```

The next part sets up the data frames to draw the map of Illinois & it's counties. I used [this](https://eriqande.github.io/rep-res-web/lectures/making-maps-with-R.html) as a guideline to figure out how to do this.

```{r state plot df}
states <- map_data("state")
il_df <- subset(states, region == "illinois")
counties <- map_data("county")
il_county <- subset(counties, region == "illinois")
```
Let's start by plotting just the state of Illinois & the county outlines.

```{r state/county plot}
ggplot(data = il_df, mapping = aes(x = long, y = lat, group = group)) +
  coord_fixed(1.3) +
  geom_polygon(color = "black", fill = "gray") +
  geom_polygon(data = il_county, fill = NA, color = "white") +
  geom_polygon(color = "black", fill = NA)
```

Looks good! Next let's plot a point for each of our airports. We add geom_point(). We have to specify the data frame that our data comes from using "data =". Then to plot the actual points we use the aes() call. I also added "size =". I used yearly_rank to determine the size of the point.


```{r all plot}
ggplot(data = il_df, aes(x = long, y = lat, group = group)) +
  coord_fixed(1.3) +
  geom_polygon(color = "black", fill = "gray") +
  geom_polygon(data = il_county, fill = NA, color = "white") +
  geom_polygon(color = "black", fill = NA) +
  geom_point(data = IL_df, aes(x = lng, y = lat, group = group, size = yearly_rank))
```

This is a good starting point! Now let's clean it up. I want to start by removing the background panel, grid lines, axis tick marks, axis labels, & axis titles. All of these can be done from the theme() call.

```{r theme change plot}
ggplot(data = il_df, aes(x = long, y = lat, group = group)) +
  coord_fixed(1.3) +
  geom_polygon(color = "black", fill = "gray") +
  geom_polygon(data = il_county, fill = NA, color = "white") +
  geom_polygon(color = "black", fill = NA) +
  geom_point(data = IL_df, aes(x = lng, y = lat, group = group, size = yearly_rank)) +
  theme(panel.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank())
```

Next thing to work on is the point size. We want the smaller values (higher ranks) to have a larger point size than the higher values (lower rank). To do this, we can take the inverse of the yearly_rank. We do this by using "1/yearly_rank" in the size call.

```{r change point size plot}
ggplot(data = il_df, aes(x = long, y = lat, group = group)) +
  coord_fixed(1.3) +
  geom_polygon(color = "black", fill = "gray") +
  geom_polygon(data = il_county, fill = NA, color = "white") +
  geom_polygon(color = "black", fill = NA) +
  geom_point(data = IL_df, aes(x = lng, y = lat, group = group, size = 1/yearly_rank)) +
  theme(panel.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank())
```

This looks better. Now let's play with the color. First, I changed the background color of the entire state to "navyblue". I did this in the first geom_polygon() call. I also changed the dots to an orange color (Go Bears!). I removed the third geom_polygon call because it was adding an black outline to the state portion. It didn't work with my color theme.

```{r change color plot}
ggplot(data = il_df, aes(x = long, y = lat, group = group)) +
  coord_fixed(1.3) +
  geom_polygon(color = "black", fill = "navyblue") +
  geom_polygon(data = il_county, fill = NA, color = "white") +
  geom_point(data = IL_df, aes(x = lng, y = lat, group = group, size = 1/yearly_rank), color = "#FF8300") +
  theme(panel.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank())
```

Next, I'm going to remove the legend using the theme() call. I'm also going to play with the point size. I'd like to see a bit more variation in the smaller points. This can be changed through the scale_size(). You can play around with the range to get the sizes you'd like.

```{r final plot}
ggplot(data = il_df, aes(x = long, y = lat, group = group)) +
  coord_fixed(1.3) +
  geom_polygon(color = "black", fill = "navyblue") +
  geom_polygon(data = il_county, fill = NA, color = "white") +
  geom_point(data = IL_df, aes(x = lng, y = lat, group = group, size = 1/yearly_rank), color = "#FF8300") +
  theme(legend.position = "none",
    panel.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank()) +
  scale_size(range = c(2, 10))
```

Looks good! Any comments or questions, you can reach out to me via [Twitter](https://twitter.com/sapo83).
