---
output: rmarkdown::github_document
---

### #TidyTuesday 9-25-18 Week 26

This week's data explores global invasion threat & cost. The data can be found [here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2018-09-25). I used the tables in the "raw" folder.

This week, we're going to do a Cleveland dot plot!

First, let's load the libraries we'll be using & read in our data. We'll use head() to take a quick peek at the data.
```{r}
library(tidyverse)
library(reshape2)

data <- read_csv("table1.csv")

head(data)
```

We can see from the head() function that there are three sets of data per row. What w're going to do is split the entire data frame into three smaller ones. The first data frame (df1) is the first three columns of the original data frame. The second data frame (df2) is the 4th through 6th column. The third data frame (df3) is the 7th through 9th column.

Then we use rename() to name the columns in the seocnd & third data frames to match the first data frame. Last of all, we combine all three data frames using rbind.

We use head() to give our new data frame a quick once over.

```{r}

df1 <- data %>%
  select(Rank, Country, OTt)

df2 <- data %>%
  select(Rank_1, Country_1, OTt_1) %>%
  rename(Rank = Rank_1, Country = Country_1, OTt = OTt_1)

df3 <- data %>%
  select(Rank_2, Country_2, OTt_2) %>%
  rename(Rank = Rank_2, Country = Country_2, OTt = OTt_2)

table1_df <- rbind(df1, df2, df3)

head(table1_df)
```

Another way to make sure that things worked how we wanted is by using nrow(). nrow() returns the number of rows in a tibble. ncol() returns the number of columns in a tibble.

Here, we can use nrow() on the three smaller data frames. Then we use nrow() on the combined data frame. The number of rows in the three smaller data frames should equal the number of rows in the combined data frame.

```{r}
nrow(df1)
nrow(df2)
nrow(df3)

nrow(table1_df)
```

42 + 42 + 42 = 126. Everything checks out so far!

Let's move onto table 3. We'll read in the data using read_csv() & then use head to look at it.

```{r}
data3 <- read_csv("table3.csv")

head(data3)
```

Well that looks... messy. It looks like there are actually two header rows here. Let's read in our data a little differently to take care of that. We can add a skip option to our read_csv() call. Here we specify "skip = 2". We'll also use "col_names = FALSE". This tells read_csv() that there are no column names.

```{r}
data3 <- read_csv("table3.csv", skip = 2, col_names = FALSE)

head(data3)
```

We got rid of the two header rows but not we have some columns that are all NAs. We'll use select() to get rid of those. By adding a minus sign ("-") before a column name in select(), we can remove that column from the data frame.

```{r}

newdata3 <- data3 %>%
  select(-X2, -X7, -X12)

head(newdata3)
```

Our next problem is that the rank & the country are in the same column. We can use separate() to fix this problem. We'll start by separating the first column. There are a couple parts to the separate() function.

  * specify the column name
  * specify separator using "sep = "
  * into takes a vector with which you can name the new columns created
  * extra = "merge" tells R to keep the extra values in the second column created.

```{r}

newdata3_1 <- newdata3 %>%
  separate(X1, sep = " ", into = c("Rank", "Country"), extra = "merge") %>%
  separate(X6, sep = " ", into = c("Rank_1", "Country_1"), extra = "merge") %>%
  separate(X11, sep = " ", into = c("Rank_2", "Country_2"), extra = "merge")

head(newdata3_1)
```

Now, we are interested in the columns that represent total invasion cost as a proportion of mean GDP. We'll use select() to keep only the columns we are interested in. To maintain the order of our columns, we going to use a minus sign (-) in our select statement to remove the columns we no longer want.

Also, let's rename the unnamed columns we are keeping. We can use rename() to do this.

```{r}

newdata3_1_filter <- newdata3_1 %>%
  select(-X3, -X4, -X8, -X9, -X13, -X14) %>%
  rename(TICt = X5, TICt_1 = X10, TICt_2 = X15)

head(newdata3_1_filter)
```

Now we're going to split our data frame like we did the first time & then bind it back together.

```{r}
df3_1 <- newdata3_1_filter %>%
  select(Rank, Country, TICt)

df3_2 <- newdata3_1_filter %>%
  select(Rank_1, Country_1, TICt_1) %>%
  rename(Rank = Rank_1, Country = Country_1, TICt = TICt_1)

df3_3 <- newdata3_1_filter %>%
  select(Rank_2, Country_2, TICt_2) %>%
  rename(Rank = Rank_2, Country = Country_2, TICt = TICt_2)

table3_df <- rbind(df3_1, df3_2, df3_3)

head(table3_df)
```

There are 42 countries in this list. I'm going to filter each table by the top 10 values. Then I'll use select to get keep only the "Country" column. I'll then combine these two lists & use distinct() to make a unique list.

```{r}
tbl1_10 <- table1_df %>%
  top_n(10, OTt) %>%
  select(Country)

head(tbl1_10)

tbl3_10 <- table3_df %>%
  top_n(10, TICt) %>%
  select(Country)

head(tbl3_10)

final_list <- rbind(tbl1_10, tbl3_10) %>%
  distinct()

```

Last of all, we're going to combine our two data frames (table1_df & table3_df) so we can start making our plot. First, we use full_join() to combine the tables. Then we use filter() to keep only the countries in our final list. We can use "%in%" to select only the countries that appear in our list.

We're also going to go ahead & drop our rank columns because we won't be needing them for the final plot. Finally, we're going to melt the data frame to make it easier to plot.

```{r}

plotdf <- full_join(table1_df, table3_df, by = "Country") %>%
  filter(Country %in% final_list$Country) %>%
  select(-contains("Rank")) %>%
  melt()

head(plotdf)
```

Let's start plotting our figure! The hallmark of a Cleveland dot plot is two values (points) connected by a line. We use geom_point() to plot each value. We use variable to color the values by group. Then we use geom_line() to connect our two points together.

```{r}

ggplot(plotdf) +
  geom_point(aes(x = value, y = Country, color = variable)) +
  geom_line(aes(x = value, y = Country))
```

Looks good so far! First, let's change up the colors & make the points bigger

```{r}

ggplot(plotdf) +
  geom_point(aes(x = value, y = Country, color = variable, size = 3)) +
  geom_line(aes(x = value, y = Country)) +
  scale_color_manual(values = c("#da532c", "#9f00a7"))
```

I'm liking that. I want to make the plot a bit cleaner by removing the x-axis. In lieu of the x-axis, I'm going to put labels next to each point with their values. I added the theme() call. Inside the theme() call, I used "axis.title" to remove the titles of both axes. I used "axis.ticks" to remove the tick marks from both axes. I used "axis.text.x" to remove the tick mark labels. I set all three of these elements to "element_blank()". Lastly, I removed the background using "panel.background".

```{r}

ggplot(plotdf) +
  geom_point(aes(x = value, y = Country, color = variable, size = 3)) +
  geom_line(aes(x = value, y = Country)) +
  scale_color_manual(values = c("#da532c", "#9f00a7")) +
  theme(axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_blank(),
    panel.background = element_blank())
```

Now, I want to add a label next to each of the points. I want the TICt labels to be on the left side of the point & I want the OTt labels to be on the right side of the point. The key here is to use two separate geom_text() calls. In the geom_text() call, you can specify your data & filter it to only include the values you want. Example: In the first geom_text() call, I specificed "plotdf" as my data source & then filtered it by "TICt" to only label those points. I also used hjust to determine where the points were plotted. I repeated this procedure with the "OTt" labels.

```{r}

ggplot(plotdf) +
  geom_point(aes(x = value, y = Country, color = variable, size = 3)) +
  geom_line(aes(x = value, y = Country)) +
  geom_text(data = plotdf %>%
    filter(variable == "TICt"), aes(x = value, y = Country, label = value), hjust = 1.25) +
  geom_text(data = plotdf %>%
    filter(variable == "OTt"), aes(x = value, y = Country, label = value), hjust = -.25) +
  scale_color_manual(values = c("#da532c", "#9f00a7")) +
  theme(axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_blank(),
    panel.background = element_blank())
```

I like the labels but we can't see some of them! We can widen the x-axis to fix this issue. I used xlim() to change the limits on the x-axis. I added 0.1 in each direction to make enough room for the labels.

I also bumped up the font size on the y-axis labels. I did this using "axis.text.y" in the theme() call.

```{r}
ggplot(plotdf) +
  geom_point(aes(x = value, y = Country, color = variable, size = 3)) +
  geom_line(aes(x = value, y = Country)) +
  geom_text(data = plotdf %>%
    filter(variable == "TICt"), aes(x = value, y = Country, label = value), hjust = 1.25) +
  geom_text(data = plotdf %>%
    filter(variable == "OTt"), aes(x = value, y = Country, label = value), hjust = -.25) +
  xlim(-0.1,1.1) +
  scale_color_manual(values = c("#da532c", "#9f00a7")) +
  theme(axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_blank(),
    panel.background = element_blank(),
    axis.text.y = element_text(size = 15))
```

Last thing I want to fix is the legend. First I want to move the legend to the bottom of the plot. This can be done inside the theme() call using legend.position = "bottom". I removed the legend title using "legend.title". Then I removed the size legend using scale_size(). I changed the labels for the legend inside the scale_color_manual() call. I changed the dot size inside the guides() call.

```{r}
ggplot(plotdf) +
  geom_point(aes(x = value, y = Country, color = variable, size = 3)) +
  geom_line(aes(x = value, y = Country)) +
  geom_text(data = plotdf %>%
    filter(variable == "TICt"), aes(x = value, y = Country, label = value), hjust = 1.25) +
  geom_text(data = plotdf %>%
    filter(variable == "OTt"), aes(x = value, y = Country, label = value), hjust = -.25) +
  xlim(-0.1,1.1) +
  scale_color_manual(values = c("#da532c", "#9f00a7"), labels = c("Overall Invasion Threat (OTt)", "Invasion Cost (TICt) as a Proportion of GDP")) +
  scale_size(guide = "none") +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  theme(legend.position = "bottom",
    legend.title = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_blank(),
    panel.background = element_blank(),
    axis.text.y = element_text(size = 15))
```

Looks great! If you have any questions or comments, feel feel to get in touch with me via [Twitter](https://twitter.com/sapo83).
